# Reproducible Research: Peer Assessment 1

### After forking GitHub Repository needed to unzip file

Used the command "unzip("activity.zip")"

## Loading and preprocessing the data

### Load and display head and tail of data
```{r getdata}
actdata <- read.csv("activity.csv")
head(actdata)
tail(actdata)
```

### Check data types
```{r checkdata}
str(actdata)
```

### Change date data from factor to date type
```{r changedata}
actdata$date <- as.Date(actdata$date)
actdata$datetime <- paste(actdata[,2],as.integer(actdata[,3]/100),actdata[,3]-100*as.integer(actdata[,3]/100))
actdata$datetime <- strptime(actdata$datetime,"%Y-%m-%d %H %M")
str(actdata)
```

## What is mean total number of steps taken per day?

### Compute the Total number of steps taken each day

```{r stepsperday}
stepsperday <- tapply(actdata$steps,actdata$date,sum,na.rm=TRUE)
hist(stepsperday)
```

###The mean and median are:
```{r meanmedian}
paste("The mean is ",mean(stepsperday), " steps per day")  
paste("The median is ",median(stepsperday),"steps per day")  
```
## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
