# Reproducible Research: Peer Assessment 1

### After forking GitHub Repository needed to unzip file

Used the command "unzip("activity.zip")"

## Loading and preprocessing the data

### Load and display head and tail of data
```{r getdata}
actdata <- read.csv("activity.csv")
head(actdata)
tail(actdata)
```

### Check data types
```{r checkdata}
str(actdata)
```

### Change date data from factor to date type
```{r changedata}
actdata$date <- as.Date(actdata$date)
actdata$datetime <- paste(actdata[,2],as.integer(actdata[,3]/100),actdata[,3]-100*as.integer(actdata[,3]/100))
actdata$datetime <- strptime(actdata$datetime,"%Y-%m-%d %H %M")
str(actdata)
```

## What is mean total number of steps taken per day?

### Compute the Total number of steps taken each day

```{r stepsperday}
stepsperday <- tapply(actdata$steps,actdata$date,sum,na.rm=TRUE)
hist(stepsperday)
```

###The mean and median are:
```{r meanmedian}
paste("The mean is ",mean(stepsperday), " steps per day")  
paste("The median is ",median(stepsperday),"steps per day")  
```
## What is the average daily activity pattern?

### Compute Total steps for each 5 min interval across days then average

```{r stepsperinterval}
stepsperinterval <- aggregate(steps~interval,actdata,sum)
stepsperinterval$steps<-stepsperinterval$steps/length(unique(actdata$date))
tail(stepsperinterval)
```

### Plot average over days of steps per interval

```{r plotintervals}
plot(stepsperinterval$interval,stepsperinterval$steps,type="l")
```

### Compute Interval of max average steps

```{r maxinterval}
paste("The interval with the max steps is ",stepsperinterval$interval[which.max(stepsperinterval$steps)])
```

## Imputing missing values

###Get Total number of NA values in steps
```{r getNA}
paste("The Total number if NA Values is " ,sum(is.na(actdata$steps)))
```

### Strategy for imputing missing values

How are NA values distributed?
```{r NADistribution}
nabydates<-tapply(is.na(actdata$steps),actdata$date,sum)
nabydates
```

Since all NA are confined to single dates and all values are missing from each of those dates it might make sense to use the mean from each interval over all the other dates
The mean should be taken though over just the dates without NA values

```{r NewMeans}
stepsperinterval2 <- aggregate(steps~interval,actdata,sum)
stepsperinterval2$steps<-stepsperinterval2$steps/length(nabydates[nabydates==0])
tail(stepsperinterval2)
```


```{r Impute}
nactdata<-actdata
sum(is.na(nactdata$steps))
nabydates<-tapply(is.na(actdata$steps),actdata$date,sum)
for(i in names(nabydates[nabydates!=0]))
     nactdata[nactdata$date==i,"steps"]<-stepsperinterval2$steps
head(nactdata)
tail(nactdata)
```

### recompute means with new data frame
```{r newmeans}
stepsperinterval <- aggregate(steps~interval,nactdata,sum)
stepsperinterval$steps<-stepsperinterval$steps/length(unique(nactdata$date))
tail(stepsperinterval)
```

The new means are different from the original computations but are consistent with the means of the old data adjusting the divisor for only the days with data
The total steps per day is not impacted since NA values were ignored for those totals
```{r nstepsperday}
nstepsperday <- tapply(actdata$steps,nactdata$date,sum,na.rm=TRUE)
hist(nstepsperday)
```


## Are there differences in activity patterns between weekdays and weekends?
